---
title: "Visualize"
format: gfm
editor: visual
---

# Climate Futures classification

This notebook executes methods to classify a climate scenario data ensemble into the National Park Service's *Climate Futures* framework. The Climate Futures framework is a storyline approach for climate scenario planning under uncertainty. In reclassifies a CMIP ensemble of climate models/ scenario combinations into new scenario, based on their actual climate inpact at a certain location.

```{r}
source('visualize.R')

plot_climate_futures('jotr')
plot_climate_futures_ensemble(park = c('jotr'))
```

├── `src` 📁 contain all the functions needed to create the data:

-   `configs.py` 📄 defined global variables

-   `dataLoader.py` 📄 is a class with methods to load various data sets. So far ISIMIP 3b, and CalAdapt (CMIP6 downscaled with LOCA2-hyrbrid) have been implemented. If you are interested in using a different data source, you will need to add a method to the `dataLoader()` class

-   `climateFutures.py` 📄 contains all the functions needed to classify scenarios and save output. If your `dataLoader()` method is written correctly, it should be able to work with any newly added dataset.

├── `outputs` 📁 contains classified output files

├── `visualize.R` 📄 defines all functions needed for the plots.

## Additional Plots

To check which model/scenario combination got classified into which climate future run

```{r}
plot_model_overview(park = 'jotr')
```

To understand, if two national parks differ in their classification run

```{r}
combine_two_parks(parks = c('jotr', 'deva'))
```

To plot time series for multiple parks, run

```{r}
plot_climate_futures_ensemble(park = c('jotr', 'deva'))
```

------------------------------------------------------------------------

This current version assumed that input climate data lives local - To download Caladapt, use `intake_caladapt.ipynb`. - To download ISIMIP 3b, use see the `isimip` folder

Note that our classification orientates itself from the original *Climate Futures* methodology, but does not completely reproduce it. Important differences: - Orignal CF uses MACA-downscaled CMIP5, we use LOCA2-downscaled CMIP6 for our main applications - Orignal CF uses observations as the baseline for anomaly calculations, we use historical simulations of the model of questions

More information on the original *Climate Futures* methodology can be found here: https://www.nps.gov/subjects/climatechange/climatefutures.htm
