---
format: gfm
editor: visual
execute:
  echo: true
  warning: false
  message: false
fig-width: 8
fig-height: 5
---

# Climate Futures classification

This notebook executes methods to classify a climate scenario data ensemble into the National Park Service's *Climate Futures* framework. The Climate Futures framework is a storyline approach for climate scenario planning under uncertainty. It reclassifies a CMIP ensemble of climate models/scenario combinations into new scenarios, aimed at enhancing the actual climate change signal at a certain location.

Lets look at an example from Joshua Tree National Park:

```{r}
#| label: setup
#| output: false
source('visualize.R')
```

```{r}
#| label: overview-jotr
plot_climate_futures('jotr')
```

```{r}
#| label: ensemble-jotr
plot_climate_futures_ensemble(parks = c('jotr'))
```

-   `src` 📁 contains all the functions needed to create the data:

    -   `config.py` 📄 defines global variables

    -   `dataLoader.py` 📄 is a class with methods to load various data sets. So far ISIMIP 3b, and CalAdapt (CMIP6 downscaled with LOCA2-hybrid) have been implemented. If you are interested in using a different data source, you will need to add a method to the `dataLoader()` class

    -   `climateFutures.py` 📄 contains all the functions needed to classify scenarios and save output. If your `dataLoader()` method is written correctly, it should be able to work with any newly added dataset.

-   `outputs` 📁 contains classified output files

-   `visualize.R` 📄 defines all functions needed for the plots.

## Additional Plots

To check which model/scenario combination got classified into which climate future run

```{r}
#| label: model-overview
plot_model_overview(park = 'jotr')
```

To understand, if two national parks differ in their classification run

```{r}
#| label: two-parks
combine_two_parks(parks = c('jotr', 'deva'))
```

To plot time series for multiple parks, run

```{r}
#| label: ensemble-two-parks
#| fig-height: 7
plot_climate_futures_ensemble(parks = c('jotr', 'deva'))
```

------------------------------------------------------------------------

This current version assumes that input climate data lives local:

-   To download CalAdapt, use `intake_caladapt.ipynb`
-   To download ISIMIP 3b, see the `isimip` folder

Note that our classification orientates itself from the original *Climate Futures* methodology, but does not completely reproduce it. Important differences:

-   Original CF uses MACA-downscaled CMIP5, we use LOCA2-downscaled CMIP6 for our main applications
-   Original CF uses observations as the baseline for anomaly calculations, we use historical simulations of the model of questions

More information on the original *Climate Futures* methodology can be found [here](https://www.nps.gov/subjects/climatechange/climatefutures.htm).
